---
format:
  html:
    theme: cosmo
---

## Objectives {.unnumbered}

- Have access to `R` through `RStudio`, `VS Code` or another GUI.
- Have basic access to the labs via GitHub classroom.
- Perform exploratory data analysis of some sample datasets.


## Quarto and GitHub Classroom

Most of the labs will be broken down into multiple tasks for you to complete.
We will extensively use **GitHub Classroom** to complete the labs.
We will talk about `git` and GitHub until Week 3, but for now you need to at least know how to get access to the lab files, so there are two commands you'll need to know.

Most of the code we write will be in `quarto`.
Quarto is a superset of `markdown`, that allows us to write documents made purely of plain text, and **compile** or **knit** them into formatted output, such as HTML or PDF.
Importantly, we can **embed `R` code** in `quarto` documents, have `quarto` run the code and embed the output of the code into the report.
This is important for reproducibility; the document we write contains all the documentation needed to explain the problem, the code to produce outputs and the interpretation of the outputs, in one place.

### Task 1

::: {.callout-tip}
The labs will contain some tasks for you to complete yourself. These will be checklists, as below.

Sometimes there will be areas for you to edit. These will have _marked_ text that [looks like this]{.mark}.
:::

- [x] Check items off as you complete them.
- [x] Create your personal student repository by following the instructions at the link above.
- [x] Clone the repository, containing this file, to your local machine.
- [x] Open these instructions in Visual Studio Code or RStudio.
- [x] Fill out your name in the box below.

::: {.callout-note}
Hi, my name is [Yuyang Jin]{.mark}.
:::

- [x] Knit the document to check that the output looks okay.
- [x] Save, and push the output back up to GitHub. (Hint: see the instructions [here](https://soton-fds.github.io/github-tutorial.html))
- [x] Check that you can see the modified version in your personal GitHub repo.

## Palmer Penguins in `R`

We will be using the Palmer Penguins dataset, a dataset that is an alternative to the classic Iris dataset.
It contains real observations of three penguin species - Adélie, Chinstrap and Gentoo - collected from the Palmer Station in Antarctica by researchers studying penguin ecology.
The dataset includes:

- Biological measurements (e.g. flipper length, body mass, bill depth)
- Metadata (species, island, sex, year)

This dataset was curated and made accessible by Allison Horst and collaborators, and has become popular in both R and Python data science communities.

The next code block loads and processes the dataset, stored in `penguins.csv` in the repository.

::: {.callout-tip}
If you look at the `.qmd` file in RStudio or VS Code, you will see the code appears surrounded by some backtick characters (\`).
These tell `quarto` that what is contained between the backticks is code.
The text after the backticks (`{r}`) tells us that this is `R` code.
When `quarto` compiles this document to HTML or PDF, the code is run, and the output is placed into the document.
:::

```{r}
# load tidyverse
library(tidyverse)
# read data from penguins.csv then store the result in penguins
penguins <- read_csv("data/penguins.csv")
# show the first few rows of the data
head(penguins)
# summarized basic statistical characteristic of penguin data
summary(penguins)
# plot a scatter graph where x and y represent flipper length and body mass respectively, color by species
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  # draw the observations as points
  geom_point() +
  # give the graph a clean visual style
  theme_minimal()
```
`The scatter plot above represent "Adelie" "Chinstrap" and "Gentoo" in red green and blue respectively. "Adelie" and "Chrinstrap" have simmilar body mass and flipper length. The body mass for most "Adelie" is in the range of 2800 g to 4800g, almost same with "Chrinstrap"; the flipper lenth for "Adelie" is in the range of 172mm to 210mm, and this range us slightly lower than "Chrinstrap". For "Genntoo", it have havier mass and larger length compare with the other two. Overall, the plot shows a positive correlation, in this case which means the penguin with longger flipper length generally tend to have higher body mass .`

### Tricks for Reproducibility

The code above is **not very well written**; it doesn't explain what is happening on each line. This is bad for reproducibility. Another scientist reading your code might not understand what is happening!

To improve this we can add *comments* to the code.
Comments are basically text that is ignored by the language, so that we can annotate it to explain it to other readers.

```{r}
# load tidyverse
library(tidyverse)
# read data from penguins.csv then store the result in penguins
penguins <- read_csv("data/penguins.csv")
# show the first few rows of the data
head(penguins)
# summarized basic statistical characteristic of penguin data
summary(penguins)
# plot a scatter graph where x and y represent flipper length and body mass respectively, color by sex
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = sex)) +
  # draw the observations as points
  geom_point() +
  # give the graph a clean visual style
  theme_minimal()

# This is what a comment in R looks like
print("Hello world!")  # We can also add comments at the end of lines.
```

### Task 2 

- [x] Add comments before each line of code in the blocks above, explaining what is happening on each line.
- [x] Add some interpretation of the plot (in `markdown`, not as comments), after the code block.
- [x] Create a second plot, below, that colours the dots by sex instead of by species.

If you don't know how to do something in `R` you have a few options:

- Check the documentation, e.g. by using `?ggplot` or similar.
- Use a search engine to find an answer.
- Ask a Generative AI tool. Most of them are pretty good at coding.

The next task asks you to *extend* what we have already done using new functionality of `ggplot`.

```{r}
# load tidyverse
library(tidyverse)
# read data from penguins.csv then store the result in penguins
penguins <- read_csv("data/penguins.csv")
# show the first few rows of the data
head(penguins)
# summarized basic statistical characteristic of penguin data
summary(penguins)
# plot a scatter graph where x and y represent flipper length and body mass respectively, color by species shape by sex
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species, shape = sex)) +
  # draw the observations as points
  geom_point() +
  # give the graph a clean visual style
  theme_minimal()
```

- [x] Create a third plot that uses colour for species and changes the *shape* of the points based on the sex.
- [x] Commit and push your solutions to GitHub with an informative message.
- [x] Make sure your changes are visible on GitHub.

From now on I won't remind you to do this - you should get into the habit of doing it whenever you reach a natural point to commit!

### Task 3

Produce plots to answer the following research questions:

- [x] On which island are we most likely to find which penguins?

```{r}
# load tidyverse
library(tidyverse)
# read data from penguins.csv then store the result in penguins
penguins <- read_csv("data/penguins.csv")
# show the first few rows of the data
head(penguins)
# summarized basic statistical characteristic of penguin data
summary(penguins)
# plot a bar chart x and y represent island and count respectively 
ggplot(penguins, aes(x = island, fill = species)) +
  # draw the observations as bar
  geom_bar(position = "dodge") +
  # give the graph a clean visual style
  theme_minimal()
```
- [ ] What is the most useful measurement or measurements to distinguish different species of penguin?



- [ ] Do the measurements given vary significantly with the penguin's genders?



